Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-01-22 manuscript.Rmd
add PSS

abstract: |
  Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. This result suggests that task-relevance is a key determinant of context-specific control.

keywords          : "Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance"
wordcount         : "4402"

@@

Another key result of this study concerned the different stimuli used as context cues. Across the three conditions, we varied the type of context image and dimensions. We manipulated the type of image presented, including both objects (identity and color dimensions) and faces (gender and racial dimensions). We also manipulated whether a single set of three repeating images were presented (object and social) or a set of non-repeating images were presented (social/non-repeating). Across all three conditions, we found no evidence that using different stimuli had an influence on the size or direction of the CSPC demonstrating generalizability of this phenomenon. Furthermore, CSPC effects were present even when using non-repeating images which suggests that context-dependency did rely on image-specific associations but higher-order, learned categorical information.

>>>>

abstract: "Evidence across a wide variety of attention paradigms shows that environmental cues can trigger adjustments to ongoing priorities for attending to relevant and irrelevant information. This context-specific control over attention suggests that cognitive control can be both automatic and flexible. For instance, in selective attention tasks, congruency effects are larger for items that appear in a context associated with infrequent conflict than in a context associated with frequent conflict. Since the to-be-presented context cannot be predicted or prepared for in advance, attention is assumed to be rapidly updated on-the-fly, triggered by the currently presented context. Context-specific control exemplifies how learning and memory processes can influence attention to enable cognitive flexibility. However, what determines the use of previously learned associations still remains unclear. In the current study, we examined whether task-relevance would influence the learning and use of context cues in a flanker task. Using a secondary counting task, context dimensions associated with differing levels of conflict were made task-relevant or -irrelevant across the experiment. In short, we found that making new contextual information task-relevant caused participants to ignore a previously learned context-attention association and adopt a new context-specific control strategy; all without changing the experimental stimuli. This result suggests that task-relevance is a key determinant of context-specific control."

keywords          : "Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance"
wordcount         : "4024 \\vspace{2mm} \\par \\textit{Public Significance Statement:} Contextual cues have been shown to automatically trigger adjustments in selective attention independent of awareness and intention. Here, we find the task-relevance plays an important role in determining which context cues are used to direct attention. These findings contribute to a better understanding of how context-dependency might occur in more complex environments and more generally, how learning and memory processes enable flexible control over attention."

@@

Another key result of this study concerned the different stimuli used as context cues. Across the three conditions, we varied the type of context image and dimensions. We manipulated the type of image presented, including both objects (identity and color dimensions) and faces (gender and racial dimensions). We also manipulated whether a single set of three repeating images were presented (object and social) or a set of non-repeating images were presented (social/non-repeating). Across all three conditions, we found no evidence that using different stimuli had an influence on the size or direction of the CSPC demonstrating generalizability of this phenomenon. Furthermore, CSPC effects were present even when using non-repeating images which suggests that context-dependency did not rely on image-specific associations but higher-order, learned categorical information.

### 2019-01-22 manuscript.tex
add PSS

\shorttitle{Contextual recruitment of selective attention}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} The Graduate Center of the City University of New York\\\textsuperscript{2} Brooklyn College of the City University of New York}
\keywords{Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance\newline\indent Word count: 4402}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
Evidence across a wide variety of attention paradigms shows that
environmental cues can trigger adjustments to ongoing priorities for
attending to relevant and irrelevant information. This context-specific
control over attention suggests that cognitive control can be both
automatic and flexible. For instance, in selective attention tasks,
congruency effects are larger for items that appear in a context
associated with infrequent conflict than in a context associated with
frequent conflict. Since the to-be-presented context cannot be predicted
or prepared for in advance, attention is assumed to be rapidly updated
on-the-fly, triggered by the currently presented context.
Context-specific control exemplifies how learning and memory processes
can influence attention to enable cognitive flexibility. However, what
determines the use of previously learned associations still remains
unclear. In the current study, we examined whether task-relevance would
influence the learning and use of context cues in a flanker task. Using
a secondary counting task, context dimensions associated with differing
levels of conflict were made task-relevant or -irrelevant across the
experiment. In short, we found that making new contextual information
task-relevant caused participants to ignore a previously learned
context-attention association and adopt a new context-specific control
strategy; all without changing the experimental stimuli. This result
suggests that task-relevance is a key determinant of context-specific
control.

}

>>>>

\shorttitle{Contextual recruitment of selective attention}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} The Graduate Center of the City University of New York\\\textsuperscript{2} Brooklyn College of the City University of New York}
\keywords{Attention, Memory, Cognitive Control, Conflict Adaptation, Context-specific, Task Relevance\newline\indent Word count: 4024 \vspace{2mm} \par \textit{Public Significance Statement:} Contextual cues have been shown to automatically trigger adjustments in selective attention independent of awareness and intention. Here, we find the task-relevance plays an important role in determining which context cues are used to direct attention. These findings contribute to a better understanding of how context-dependency might occur in more complex environments and more generally, how learning and memory processes enable flexible control over attention.}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
Evidence across a wide variety of attention paradigms shows that
environmental cues can trigger adjustments to ongoing priorities for
attending to relevant and irrelevant information. This context-specific
control over attention suggests that cognitive control can be both
automatic and flexible. For instance, in selective attention tasks,
congruency effects are larger for items that appear in a context
associated with infrequent conflict than in a context associated with
frequent conflict. Since the to-be-presented context cannot be predicted
or prepared for in advance, attention is assumed to be rapidly updated
on-the-fly, triggered by the currently presented context.
Context-specific control exemplifies how learning and memory processes
can influence attention to enable cognitive flexibility. However, what
determines the use of previously learned associations still remains
unclear. In the current study, we examined whether task-relevance would
influence the learning and use of context cues in a flanker task. Using
a secondary counting task, context dimensions associated with differing
levels of conflict were made task-relevant or -irrelevant across the
experiment. In short, we found that making new contextual information
task-relevant caused participants to ignore a previously learned
context-attention association and adopt a new context-specific control
strategy; all without changing the experimental stimuli. This result
suggests that task-relevance is a key determinant of context-specific
control. 
}
